---
title: "task1"
author: "Linas Syvis"
date: "Thursday, February 11, 2016"
output: html_document
---

This is Task1 for Practical Econometrics II course made by Linas Syvis. The code was taken from the source: http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/ and the whole task was made by an example from: http://shiny.rstudio.com/articles/rmarkdown.html


These two data sets will be used to generate the graphs below.
```{r}
# Two variables
df <- read.table(header=TRUE, text="
 cond yval
    A 2
    B 2.5
    C 1.6
    Z 3.8
")

# Three variables
df2 <- read.table(header=TRUE, text="
 cond1 cond2 yval
    A      I 2
    A      J 2.5
    A      K 1.6
    B      I 2.2
    B      J 2.4
    B      Z 1.2
    C      I 1.7
    C      J 2.3
    C      K 1.9
")
```
Instead of changing colors globally, you can map variables to colors – in other words, make the color conditional on a variable, by putting it inside an aes() statement.
You will also need to download and install a package called "ggplot2".
```{r}
install.packages('ggplot2')
library("ggplot2")

# Bars: x and fill both depend on cond2
ggplot(df, aes(x=cond, y=yval, fill=cond)) + geom_bar(stat="identity")

# Bars with other dataset; fill depends on cond2
ggplot(df2, aes(x=cond1, y=yval)) + 
    geom_bar(aes(fill=cond2),   # fill depends on cond2
             stat="identity",
             colour="purple",    # Purple outline for all
             position=position_dodge()) # Put bars side-by-side instead of stacked

# Lines and points; colour depends on cond2
ggplot(df2, aes(x=cond1, y=yval)) + 
    geom_line(aes(colour=cond2, group=cond2)) + # colour, group both depend on cond2
    geom_point(aes(colour=cond2),               # colour depends on cond2
               size=3)                          # larger points, different shape
## Equivalent to above; but move "colour=cond2" into the global aes() mapping
# ggplot(df2, aes(x=cond1, y=yval, colour=cond2)) + 
#    geom_line(aes(group=cond2)) +
#    geom_point(size=3)
```
